Using Stochastic Modeling to Predict Long Timescale Transport Behavior of Solutes in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2020-01-17 Ben_Manuscripts/hdphmm/Outline.tex
Starting outline of HDP-HMM paper

%MRS: better to have a title that described a finding, rather than the
%approach used.  But we may need to wait a bit further until we can decide on a finding.
%MRS: also 'stochastic modeling' is pretty vague.  should be more precise. 
\title{Time Series Modeling of Solute Transport in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane}
%MRS: we should discuss whether Chris Calderon would be on this one.  I would guess no, but I don't know for sure. 
%BJC1: Yea maybe Chris as an acknowledgement on this one. He didn't directly contribute anything other than pointing
%me towards Greg Morrin who pointed me towards useful anomalous diffusion literature.
%MRS1: I wonder if ``Times series modeling'' is better?  We should keep thinking about a better term.  Stochastic modeling is such a general term, it encompasses a lot  . . . 
\author{Benjamin J. Coscia \and Michael R. Shirts}

@@

  \section{Introduction}

  % MRS1: I would start with something more central to the content of the paper - how do solutes move in complex separations membranes?  Let people know right from the start what the paper is about.  This paragraph should still be in the intro, but start with something that captures the main question of the paper. 
  % BJC2: This paragraph will need more tweaking but is the gist of what I'm thinking
  % BJC2: I'm not sure if this is exactly what you meant though
%  Making sense of solute motion in complex separations membranes is central to 
%  understanding transport mechanisms and formulating membrane design principles.
  Mathematical descriptions of transport in complex separations membranes are a powerful 
  way to understand mechanisms and formulate membrane design principles.
  % MRS3: seems pretty good.  We can tweak later when turning into text; I don't think that it will really affect the logic of the paper. The introduction will be important here to set the stage, but probably more of that will be in the next paragraph.
  \begin{itemize}
    \item The complexity of a model generally parallels the complexity of the 
    transport mechanism at hand, as well as the desired outputs.
    \item In dense homogeneous membranes, the solution-diffusion model can extract
    diffusion and partition coefficients and has successfully predicted solute transport
    rates.
    \item Analogously, pore-flow models yield predictions of diffusion coefficients
    and solute transport rates in nanoporous membranes.~\cite{wijmans_solution-diffusion_1995}
    \item Single particle tracking allows us to further characterize more complex
    diffusive behavior.
    \item Finally, molecular simulation provides a framework for studying both single
    particle dynamics and bulk transport properties with atomic resolution.
    \item All of these approaches facilitate speculation on the molecular origins of
    separations and suggest experiments that should be run by attempting to give a more
    intuitive understanding of how solutes move as a function of their environment.
%    \item The hydrodynamic Ferry model is the simplest pore-flow model, with extensions such
%    as the Donnan-Steric Pore Model which includes extra terms for the pressure gradient
%    across the membrane and membrane charge.
%    \item Dischinger et al. created an empirical correlation to determine the
%    effect of solute physicochemical properties on solute rejection.
  \end{itemize}

%  \begin{itemize}
%    \item More recently, rapid advancements in computational power have made it feasible
%    to observe transport in simulated systems with atomic resolution. This allows us to
%    shift towards analysis of the motion of individual particles and gives an even more 
%    fundamental understanding of observed bulk properties.
%  \end{itemize}

  % BJC1: tie this paragraph into above by commenting that highly selective membranes
  % often have complex mechanistic reasons that aren't explained by simpler models
  These approaches can be especially useful for designing highly selective membranes. \\
  %BJC2: this paragraph will contain examples of highly selective separations

  %MRS3: I'd omit next line (That I commented out); you talk about membranes above, so better to start the next paragraph talking about membranes.
  %\noindent Amphiphilic molecules are capable of self-assembling into ordered nanostructures.\\

  Lyotropic liquid crystals are a class of amphiphilic molecules that can be cross-linked
  into mechanically strong and highly selective membranes.
  \begin{itemize}
  	\item H\textsubscript{II} phase lyotropic liquid crystals have densely packed, uniform
	sized pores and have the potential to disrupt conventional membrane separation
	techniques by being selective based not only on size and charge, but on chemical
	functionality as well.
	\item Q\textsubscript{I} phase LLCs consist of a tortuous network of 3D interconnected
	pores. They are easier to make.
        %MRS3: probably want to put a more here about why we are studing these membranes, and what advantages they might have, and why transport in them is complicated.  
  \end{itemize}

  %MRS3: One missing connection between the above information and what is in the paper would be making the connection between MD simulations
  %and mathematical modeling, using MD to provide/justify the fuctional form and/or some of the parameters for the model, 
  %rather than just experiment.

  %MRS3: a suggested tweak in language
  %We can only learn so much from experiment. 
  There are limitations to what we can learn from experiment about the atomistic details of transport.
  MD can give us mechanistic insights with
  atomistic resolution so that we can intelligently design new membranes for 
  solute-specific separations.\\
  %BJC2: simple models from first paragraph fail to adequately describe transport.

  \noindent In our previous work, we studied the transport of 20 small polar molecules
  in an H\textsubscript{II} phase LLC membrane.
  \begin{itemize}
    \item In general, we observed subdiffusive transport behavior characterized by 
    intermittent hops separated by periods of entrapment.
    \item We identified three mechanisms responsible for the solute trapping behavior:
    entanglement among monomer tails, hydrogen bonding with monomer head groups, and
    association with sodium counter ions.
    %MRS3: point out that they were all subdiffusive (unlike water)
  \end{itemize}

  Unfortunately, the timescales that we can simulate with MD are insufficient to be
  able to make well-converged predictions of macroscopic transport properties 
  traditionally used to characterize membranes in the lab.
  \begin{itemize}
    \item However, if we use descriptive stochastic models that sufficiently capture solute
    dynamics, then we could project long timescale behavior in addition to gaining
    a deeper understanding of solute behavior on short timescales.
  \end{itemize}

  \noindent Researchers have developed a rigorous theoretical foundation for describing
  the motion of particles exhibiting non-Brownian, or anomalous, transport behavior.
  \begin{itemize}
    \item The tools introduced by fractional calculus are instrumental to this theory. 
    \item They allow us to generalize the normally linear diffusion equation to fractional
    derivative orders, providing descriptions of a much more diverse set of behavior. \cite{klages_anomalous_2008} % diverging from the relatively simple Brownian model.
%    \item They facilitate proof of the relationship between the observed growth of a
%    particle's mean squared displacement and underlying dwell times 
%    \item In the absence of an external potential or complicated boundary conditions, 
%    one can define random walk based on independent hop length and dwell time 
%    distributions and arrive at any type of diffusive behavior. 
  \end{itemize}

  % BJC1: moved from methods
  \noindent Three well-known classes of behavior leading to anomalous subdiffusion are 
  continuous time random walks (CTRW), fractional Brownian motion
  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
  \begin{itemize}
    \item The theory is frequently used to describe single particle trajectories
    \item A CTRW is characterized by a distribution of hop lengths and 
    dwell times, where each each hop is characterized by independent random draws from 
    each distribution.\cite{montroll_random_1965,morrin_three_2018}
    \item FBM is common in crowded, viscoelastic environments where each jump comes 
    from a Gaussian distribution but is anti-correlated to its previous 
    step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
    \item An RWF is imposed by a system's geometry. Systems with tortuous pathways and dead
    ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}
    \item The processes described above can happen alone or in combination. 
    \item Meroz and Sokolov suggested a decision tree for determining which subdiffusive
    mechanisms apply to a system, adapted in Figure X. 
    \item Of course, more complicated variations on these models can arise which we 
    will incorporate as needed.	  % BJC1: e.g. fractional levy motion or langevin with levy noise
  \end{itemize}

  % BJC1: Flow chart here

  % BJC2: moved this from below
  Our first modeling approach is based on the anomalous diffusion literature, applying the 
  mathematical formalism describing subdiffusive transport to observed solute 
  trajectories.
  \begin{itemize}
    \item We determine the most likely subdiffusive mechanisms at play using the 
    flow chart in figure X. 
    \item We then fit anomalous diffusion models in two ways: First, based on a
    single set of parameters. Second, we use a two-state approach incorporating
    two sets of parameters based on a solute's radial distance from the pore center. 
  \end{itemize}

  While elegant, pure anomalous diffusion theory may in some cases give only a qualitative
  understanding of transport mechanisms.
  \begin{itemize}
    \item Complex temporal relationships in between hops may have a significant effect on the average MSD
    \item Inhomogeneity of a medium or multiple types of interactions may lead to a number of
    distinct diffusive regimes. 
    \item In this case, a state based model may be useful. 
  \end{itemize}

%  \noindent We can use the observations of transport from our previous work in order to create 
%  state-based models.
%  \begin{itemize}
%  	\item We know that dynamics differ between solutes moving inside the pores and 
%  	in the tails.
%  	\item We can use the anomalous diffusion approach to model our data as a 
%  	function of radial distance from the closest pore center.
%  	\item Alternatively, we can abandon the anomalous diffusion theory and adapt the 
%  	methods of Markov state modeling.
%  \end{itemize}

  % moved parts of method to here.
  Markov state models (MSMs) are a popular class of models used to project long timescale
  behavior of molecules based on molecular simulation trajectories by identifying
  different dynamical modes and quantifying the rates of transitions between them.
  \begin{itemize}
    \item MSMs are frequently used to study systems with slow dynamics, such as protein 
    folding.~\cite{snow_how_2005,chodera_automatic_2007}
    \item Researchers typically aim to come up with a low dimensional 
    representation of the system based on features which preserve the process
    kinetics.
    \item This still often results in thousands of distinct states. 
  \end{itemize}

  %MRS1: probably want to make clear that you really are doing some
  %substantially different things in in MSM modeling than most people
  %do, by adding on modeling the z-diffusion as an auxiliary variable
  %of the state.  I think that's actually kind of innovative.

  %BJC2: moved this paragraph up from below to clarify novelty of this approach
  Our second modeling approach adapts the MSM framework to a relatively small set
  of known states based on the three observed solute trapping mechanisms.
  \begin{itemize}
    \item Standard MSMs assume first order kinetics and are typically applied 
    towards determining equilibrium populations of various state configurations.
    \item We extend the framework to include dynamics within states and
    between jumps. 
    \item That is we incorporate the magnitude and temporal dependence of the 
    time series fluctuations between each transition to a different state. 
    \item Temporal dependence has a significant influence on mean squared displacement.
  \end{itemize}

  In this work we apply the two modeling approaches in order to predict the long
  timescale behavior of the four fastest moving solutes studied in our previous work.
  \begin{itemize}
    \item Specifically, we study methanol, urea, ethylene glycol and acetic acid
    \item In addition to moving quickly, these solutes have a range of chemical
    functionality and experience each of the three trapping mechanisms to different
    degrees.
    \item It is our goal to reproduce the MSDs exhibited by each solute on MD timescales
    and then to project the long-term dynamics of each solute. 
%  	\item The third uses an infinite hidden Markov model (iHMM), an unsupervised machine
%  	learning algorithm which groups solute behavior into different dynamical states.
  	% BJC: somewhere emphasize that this last approach is powerful since it doesn't 
  	% require anything different in systems with more complex geometries (i.e. QI)
  \end{itemize}

  % BJC: I've been trying to understand the following justification a bit more. Does it
  % rely on the MSD becoming linear so we can get experimentally diffusion constants? Is
  % it just to try and reduce uncertainty in our MSD predictions? What can we learn from
  % long stochastic simulations that we can't learn from the shorter simulations.
  % BJC: To me the value in this is gaining a deeper mechanistic understanding by 
  % putting things in terms of easily interpreted equations.
  % BJC: Maybe it's just a combination of both.

  % MRS: They way to think about it, I think, is that the simplest
  % assumption is the the transport mechanism is simple diffusion;
  % uncorrelated random jumps due to molecular collisions. If you make
  % that assumption, you get MSD linear in time.  So that's IS a
  % stochastic process, but not a very complex one. To make long
  % timescale predictions, we need to increase the detail.  Is that
  % what you are referring to? Not sure if I answered the question . . . 
  % BJC1: Yea, I see what you're saying. Just not sure what else the long
  % timescale will tell us that we can't learn from short ones. Probably 
  % better to talk about this in person.

%  We can use the observed mechanisms in order to choose a stochastic model of
%  sufficient complexity One can project long timescale solute behavior by modeling transport as a
%  stochastic process.
%  %MRS: as above, simple diffusion is stochastic, just not very complex.  So we want to add appropriate complexity.
%  \begin{itemize}
%    %MRS: best talk about these mechanisms in terms of the
%    %mathematical models they capture.  But I don't know that it's
%    %necessary to review all of them.  
%  
%    %It may be a better approach to start from the features of the
%    %problem (i.e. certain mechanisms) and choose the model (of the
%    %existing choices) that is appropriate for those mechanisms.
%    %I.e. diagnose first, then choose, rather than just applying a
%    %list of models useful in other field's timeseries. Let me know if that makes sense;
%    %I think how the methods applied are selected is a key to solve the problem.
%
%    \item Various disciplines study this type of problem in a number of ways.
%    \item In the simulation literature, most start by measuring mean squared 
%    displacement (MSD), as we did previously.  % This doesn't quite fit the thesis
%    \item Due to the short time scales most people can feasibly simulate, the MSD 
%    does not give a complete picture of what happens long-term, other than hinting
%    at types of mechanisms that could lead to the curve's shape.
%    %MRS: I'd say that MSD is too simple compared to what is going on mechanistically, but it's fundamentally the same idea.  
%    \item mean first passage time?
%    \item Markov state models are popular with researchers of proteins which have 
%    many dynamical states occurring over multiple time scales.
%    \item The anomalous diffusion literature explains the shape of MSD curves based
%    on rigorous mathematical representations of transport mechanisms. This field is
%    heavily used in analysis of single particle tracking experiments.
%    \item Finally, the financial literature tackles similar problems from a more 
%    general time series perspective. Their primary focus is towards predictions of
%    stock market returns and volatility but their methods are widely applicable to
%    modeling time series that exhibit a range of behavior.
%  \end{itemize}

@@

  We ran all MD simulations and energy minimizations using GROMACS 2018. We  % needs gromacs citations
  performed all post-simulation trajectory analysis using python scripts which are available
  %MRS: eventually want to say what scripts were used to do what, but that is not at the outline stage!
  %BJC1: I usually put a table like that in the SI. Is that sufficient?
  online at \\ \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 
  10 \% water by weight. (See Figure tbd)
  \begin{itemize}
    \item Approximately one third of the water molecules occupy the tail region 
    with the rest near the pore center.
    \item We chose to study the 10 wt \% water system because solutes move 
    significantly faster than in the 5 wt \% system studied previously.
    \item Appropriate stochastic modeling requires that solutes sample the 
    accessible mechanisms with representative probability.  %BJC1: incorporated suggested phrasing
  \end{itemize}

@@

  We determined equilibration based on the solute partition between the pore and tail
  region.

  \subsection{The Anomalous Diffusion Model}\label{method:model_sFBM}

  Solutes in this system exhibit subdiffusive behavior, a type of anomalous diffusion.
  \begin{itemize}
  %MRS: you would probably want to be clear what the mechanistic reasons for for anomalous diffusion: 
  %want to talk about why, given the mechanisms you would consider. 
  %BJC1: could go through flow chart
  	\item During an anomalous diffusion process, the mean squared displacement (MSD)
  	does not grow linearly with time, rather it follows a power law of the form:
	\begin{equation} 
	\langle x^2(t) \rangle = K_{\alpha}t^\alpha
	\label{eqn:msd_form}
	\end{equation} 
	where $\alpha$ is the anomalous exponent and $K_\alpha$ is the
	generalized diffusion coefficient.
	\item A value of $\alpha < 1$ indicates a subdiffusive process, while values of
	$\alpha = 1$ and $\alpha > 0$ are characteristic of Brownian and superdiffusive
	motion respectively.
  \end{itemize}

  \noindent We analyzed both the ensemble-averaged and time-averaged MSDs
  of the simulated trajectories.
  \begin{itemize}
    %MRS: describe the physical situations in which one would want to choose each of these ways of calculating the MSD.
    %BJC1: For determining ergodicity. Ensemble for determining anomalous exponent. Not sure I'll use ensemble MSD
	\item The ensemble-averaged MSD measures the displacement of a particle from its initial
	position~\cite{meroz_toolbox_2015} and can be written as
	\begin{equation}
	\langle x^2(t) \rangle = \langle x(t) - x(0) \rangle^2
	\label{eqn:ensemble_msd}
	\end{equation}
%	\item The average MSD calculated in this way recovers the form of
%	Equation~\ref{eqn:msd_form}. % time-averaged should recover it too for FBM 
	\item The time-averaged MSD measures the displacement between all possible time lags
	and can be written as
	\begin{equation}
	\overline{x^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (x(t + \tau) - x(t))^2 dt
	\end{equation}
	where $\tau$ is the time lag and T is the length of the
	trajectory~\cite{meroz_toolbox_2015}. 
  \end{itemize}

  % BJC1: moved this to intro
%  \noindent Three common mathematical models for modeling anomalous subdiffusion
%  processes include continuous time random walks (CTRW), fractional Brownian motion
%  (FBM) and random walks on fractals (RWF).\cite{meroz_toolbox_2015}
%  \begin{itemize}
%    \item FBM is common in crowded, viscoelastic environments where each step comes 
%    from a Gaussian distribution but is anti-correlated to its previous 
%    step.~\cite{mandelbrot_fractional_1968,jeon_fractional_2010,banks_anomalous_2005}
%    \item A CTRW is characterized by a distribution of hop lengths and 
%    dwell times, where each each step is characterized by independent random draws from 
%    each distribution.\cite{montroll_random_1965,morrin_three_2018}
%    \item An RWF is imposed by a system's geometry. Systems with tortuous pathways and dead
%    ends cause anti-correlated motion.\cite{meroz_toolbox_2015,neusius_subdiffusion_2008}
%    \item The processes described above can happen alone or in combination.  	
%  \end{itemize}
  %MRS: talk about which of these should apply given the data mechanistically, as well as fit.
  %BJC1: I believe that's what I try to do in the paragraph below. I can be a little more explicit

  % BJC: Should this be in the results? Lots of figures to show in support of the claims below
  % MRS: can you provide mechanistic reasoning for the below? Then can show the claims for the reasonable possibilities.
  % BJC1: modified this paragraph to parallel the flow chart. I'm still not sure the
  % best place for this. It seems like it belongs in the discussion, but I need to 
  % justify the choice in model so I can write the methods about the chosen models
  % BJC2: Perhaps just say that we think it's sFBM and move the proof to discussion.
  \noindent We believe that solutes in the system studied here exhibit subordinated 
  fractional Brownian motion (sFBM) where the parent process is FBM and the 
  leading process is a CTRW. 
  \begin{itemize}
    \item Working down the flow chart...
  	\item We observe non-stationary $z$-coordinate traces of each solute's
  	center of mass (COM). %BJC: figure of a z-coordinate trace in main text or in supporting info
  	\item Combined with the observation that the ensemble-averaged MSD differs
  	from the time-averaged MSD, we have evidence of non-ergodicity in this system, 
  	a trait inherent to CTRWs but not FBM or RWFs.~\cite{thiel_weak_2014}  % BJC: need a figure to show this
% BJC1: following two points are weak. Linearity of CTRW doesn't necessarily hold up on short time scales.
%  	\item For a pure CTRW, the time-averaged MSD should be linear.
%  	~\cite{neusius_subdiffusion_2008,meroz_subdiffusion_2010}
%  	\item However, a typical time-averaged solute MSD is sublinear (see supporting
%  	information), which suggests that there is another underlying subdiffusive mechanism.
	\item The process is likely subordinated to another process because the hop 
	lengths recorded after each dwell period are anti-correlated.
    % BJC1: need to eliminate RWF. I think a geometric argument should be convincing enough. Our
    % system doesn't have a fractal-like structure. Monomer positions fluctuate enough that
    % solutes can access any part of the membrane over time.
  	\item Given the viscoelastic nature of the monomers in our system, we believe
  	the hop lengths can be modeled with FBM. 
  	% BJC1 : probably won't need this equation. It's really just to validate my sfbm
  	% simulation procedure. Could go in supporting information with a plot showing that
  	% my sims fit to t^alpha*beta
 	\item For subordinated FBM, it can be shown that
  	\begin{equation}
  	\langle x^2(t) \rangle \simeq t^{\alpha\beta}
  	\end{equation}
  	where $\alpha$ is the anomalous exponent characteristic of the leading CTRW process
  	and $\beta$ is the anomalous exponent characteristic of the parent FBM process. 
  \end{itemize}

  \noindent We can characterize a CTRW process using the parameters which describe its
  dwell time and hop length distribution.  % This is different from the emissions of the markov state model where every single step is used to construct the distribution. 
  \begin{itemize}
	\item We used the \texttt{ruptures} python package in order to automatically identify
	changepoints in solute trajectories.\cite{truong_ruptures:_2018} (See Supporting
	Information for more details on chosen parameters. i.e. type of cost function, 
	cost function penalty tolerance, number of dimensions used)
	\item We used the corresponding hop lengths and dwell times between break points
	to construct empirical distributions.
  \end{itemize}

  % BJC1: not sure about this anymore. Might eventually change to Levy. But in the 
  % interest of simplicity, maybe leave it at Gaussian.
  % BJC2: This will change to Levy. 
  For solutes in our system, the distribution of hop lengths appears to be
  well-represented by a Gaussian distribution.~\cite{metzler_random_2000,
  metzler_anomalous_2014,neusius_subdiffusion_2009}  % Note that this is different from the distribution of all increments (which is heavy-tailed)
  \begin{itemize}
	\item We are most interested in the standard deviation, $\sigma$, of the 
	hop length distribution.
  \end{itemize}

  \noindent The distribution of dwell times is expected to fit a power law
  distribution proportional to $t^{-1-\alpha}$.~\cite{meroz_toolbox_2015}
  \begin{itemize}
	\item Because we are limited to taking measurements at discrete values
	dictated by the output frequency of our simulation trajectories, we fit the
	empirical dwell times to a discrete power law distribution whose maximum
	likelihood $\alpha$ parameter we calculated by maximizing the following
	likelihood function: 
    \begin{equation}
      %MRS: use \ln instead of ln or \text{ln} 
	\mathcal{L}(\beta) = -n\ln \zeta(\beta, x_{min}) -
	\beta\sum_{i=1}^{n} \ln x_i 
	\label{eqn:powerlaw_likelihood}
	\end{equation}
	where $\beta = 1 + \alpha$, $x_i$ are collected dwell time data points,
	$n$ the total number of data points, and $\zeta$ is the Hurwitz zeta function
	where $x_{min}$ is the smallest measured value of
	$x_i$.~\cite{clauset_power-law_2009} 
	\item We obtained distributions of the hop length standard deviations, $\sigma$, and
	$\alpha$ using bootstrapping.\cite{efron_introduction_1994} 
  \end{itemize}

  In practical applications, the heavy tail of power law distributions can result in 
  arbitrarily long dwell times that are never observed in MD simulations. 
  \begin{itemize}
    \item We can not observe dwell times longer than the length of a single trajectory.
    \item In fact the longest dwell time observed by solutes in this study is X.
    \item In order to compare our anomalous diffusion model to MD results we need to 
    bound the dwell time distribution.
    \item We do this by adding an exponential cut-off to the power law so the dwell 
    time distribution is of the form:
    \begin{equation}
    t^{-1 - \alpha}e^{-\lambda t}
    \end{equation}
    \item We determine MLEs of $\alpha$ and $\lambda$ by maximizing the likelihood function:~\cite{clauset_power-law_2009}
    \begin{equation}
    %MRS3: I replaced ln with \ln
    \mathcal{L}(\alpha, \lambda) = n(1 - \alpha)\ln\lambda - n\ln\Gamma(1 - \alpha, x_{min}\lambda) - \alpha\sum_{i=1}^{n}\ln x_i - \lambda\sum_{i=1}^n x_i
    \end{equation}
  \end{itemize}

  \noindent FBM processes can be described using the Hurst parameter, $H$.
  \begin{itemize}
  	\item Brownian motion is recovered for $H = 0.5$
	\item The autocovariance function of hop lengths has the analytical form:~\cite{mandelbrot_fractional_1968}
    \begin{equation}
	\gamma(k) = \dfrac{1}{2}\bigg[|k-1|^{2H} - 2|k|^{2H} + |k+1|^{2H}\bigg]
	\label{eqn:autocovariance}
	\end{equation}
	where $k$ is the number of increments between hops.
	\item We obtained H by performing a least squares fit of Equation~\ref{eqn:autocovariance}
	to the empirically measured autocovariance function.
	\item We used statistical bootstrapping to generate a distribution of $H$ 
	values. %BJC: can provide more detail in supporting or just reference python script
	\item There are more complicated ways to calculate H but well-converged estimates 
	require far more data than we have access to.
  \end{itemize}

  \noindent If the distribution of hops is not Gaussian, we can model them with the
  more general class of L\'evy stable distributions.
  \begin{itemize}
    \item For independent and identically distributed (iid) random variables, the 
    generalized central limit theorem guarantees convergence of the associated 
    probability distribution function (PDF) to a L\'evy stable PDF. \cite{klages_anomalous_2008}
    % BJC: ``For sums of independent, iid random variables with proper normalization to the
%    sample size, the generalized central limit theorem guarantees the convergence of
%    the associated PDF to a L\'evy stable PDF even though the variance of these random
%    variables diverges." \cite{klages_anomalous_2008}  % BJC: reword
    \item The characteristic equation describing the Fourier transform of a L\'evy 
    stable PDF is given below:
    \begin{equation}
    p_{\alpha, \beta}(k;\mu,\sigma) =\exp\left[i\mu k - \sigma^{\alpha}|k|^{\alpha}\left(1 - i\beta\frac{k}{|k|}\omega(k, \alpha)\right)\right]
    \end{equation}
    \item Where $\omega$ is defined with something that is complicated to format and can be 
    added later if I keep these formulas in. % https://tex.stackexchange.com/questions/32140/how-to-write-a-function-piecewise-with-bracket-outside
    \item $\alpha$ is the index of stability or L\'evy index, $\beta$ is the skewness 
    parameter, $\mu$ is the shift parameter and $\sigma$ is a scale parameter.
    \item The most familiar case, and 1 of 3 that can be expressed in terms of elementary
    functions, is the Gaussian PDF ($\alpha$ = 2).
    \item We assume symmetric distributions centered about 0 implying that $\beta$ and $\mu$ are both 0.
  \end{itemize}

  \noindent If the hops are correlated, we can describe their motion as subordinated 
  fractional L\'evy motion.
  \begin{itemize}
    \item The hurst parameter can again be used to describe the degree of anitcorrelation.
    \item It has the form: %BJC2: need to see if there is an analytical form for H
    %BJC2: might need to use a different way to estimate H (which is more standard): https://pypi.org/project/hurst/
  \end{itemize}

  In general, we observe different dynamical behavior when solutes move while inside 
  the pore versus while in the tail region.
  %MRS: could present the two models sequentiually.  Talk about one COULD have multiple modes, then in results present 1, then 2. 
  \begin{itemize}
    \item This inspired two models of varying complexity.
    \item We created a simple, single mode model with a single $\alpha$, $\sigma$ and $H$
    parameter fit to each solute.
    \item Our second, two mode model assigns a set of parameters to each of 2 modes based
    on the solute's radial location.
    \item Solutes in mode 1 are in the pore region defined as less than 0.75 nm from any
    pore center and all else are in mode 2, the tail region. 
    \item We determined this cut-off and described how we calculated radial 
    distance from the pore center in our previous work~\cite{coscia_chemically_2019}
  \end{itemize}

  For the two mode model, we needed to define a probability transition matrix 
  describing transitions between the tail and pore region. \\

>>>>

%MRS: better to have the first outline shorter to iterate more quickly!
%BJC1: right, this one has been in the works for a while now. Wrote the anomalous diffusion methods a long time ago

%MRS: better to have a title that described a finding, rather than the
%approach used.  But we may need to wait a bit further until we can decide on a finding.
%MRS: also 'stochastic modeling' is pretty vague.  should be more precise. 
\title{Using Stochastic Modeling to Predict Long Timescale Transport Behavior of Solutes in an H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane}
%MRS: we should discuss whether Chris Calderon would be on this one.  I would guess no, but I don't know for sure. 
%BJC1: Yea maybe Chris as an acknowledgement on this one. He didn't directly contribute anything other than pointing
%me towards Greg Morrin who pointed me towards useful anomalous diffusion literature.
\author{Benjamin J. Coscia \and Michael R. Shirts}

@@

  \section{Introduction}
  %BJC: obviously this section will get beefed up
  We need highly selective membranes in order to perform efficient separations. \\

  \noindent Amphiphilic molecules are capable of self-assembling into ordered nanostructures.\\

  Lyotropic liquid crystals are a class of amphiphilic molecules that can be cross-linked
  into mechanically strong membranes.
  \begin{itemize}
  	\item H\textsubscript{II} phase lyotropic liquid crystals have densely packed, uniform
	sized pores and have the potential to disrupt conventional membrane separation
	techniques by being selective based not only on size and charge, but on chemical
	functionality as well.
	\item Q\textsubscript{I} phase LLCs consist of a tortuous network of 3D interconnected
	pores. They are easier to make.
  \end{itemize}

  We can only learn so much from experiment. MD can give us mechanistic insights with
  atomistic resolution so that we can intelligently design new membranes for 
  solute-specific separations.\\

  \noindent In our previous work, we studied the transport of 20 small polar molecules
  in an H\textsubscript{II} phase LLC membrane.
  \begin{itemize}
    \item In general, we observed subdiffusive transport behavior characterized by 
    intermittent hops separated by periods of entrapment.
    \item We identified three mechanisms responsible for the solute trapping behavior:
    entanglement among monomer tails, hydrogen bonding with monomer head groups, and
    association with sodium counter ions.
  \end{itemize}

  Unfortunately, the timescales that we can simulate with MD are insufficient to be
  able to make well-converged predictions of macroscopic transport properties 
  traditionally used to characterize membranes in the lab.
  \begin{itemize}
    \item However, if we use descriptive stochastic models that can capture solute
    dynamics, then we could project long timescale behavior in addition to gaining
    a deeper understanding of solute behavior on short timescales.
  \end{itemize}

  In our previous work, we designed two different approaches which used
  solute time series in order to parameterize stochastic models that could be used
  to project transport on much longer timescales.
  \begin{itemize}
  	\item Brief description of anomalous diffusion
  	\item Brief descritpion of MSDDM
  \end{itemize}

  Although both models had reasonable success at predicting solute MSDs on simulation
  timescales, they had shortcomings.
  \begin{itemize}
  	\item Why MSDDM failed
  	\item Why anomalous diffusion model could be better.
  \end{itemize}

  % BJC: Not sure whether to call it the infinite hidden markov model or hierarchical dirichlet process hidden markov model
  % The former is definitely simpler.
  In this work, we apply the infinite hidden markov model, a modeling that is agnostic
  to the source of time series data, in order to detect and parameter an unknown number
  of autoregressive modes.

@@

  We ran all MD simulations and energy minimizations using GROMACS 2018. We  % needs gromacs citations
  performed all post-simulation trajectory using python scripts which are available
  %MRS: eventually want to say what scripts were used to do what, but that is not at the outline stage!
  %BJC1: I usually put a table like that in the SI. Is that sufficient?
  online at \\ \texttt{https://github.com/shirtsgroup/LLC\_Membranes}.

@@

  We studied transport of solutes in the H\textsubscript{II} phase using an
  atomistic molecular model of four pores in a monoclinic unit cell with 
  10 \% water by weight. 
  \begin{itemize}
    \item Approximately one third of the water molecules occupy the tail region 
    with the rest near the pore center.
    \item We chose to study the 10 wt \% water system because solutes move 
    significantly faster than in the 5 wt \% system studied previously.
    \item Appropriate stochastic modeling requires that solutes sample the 
    accessible mechanisms with representative probability.  %BJC1: incorporated suggested phrasing
  \end{itemize}

@@

  % BJC: can ignore this section for now. 
  \subsection{The Infinite State Hidden Markov Model}\label{method:iHMM}  %BJC: potentially a separate paper

  Hidden Markov models (HMMs) are a useful and widely used technique
  for modeling sequences of observations where the probability of the next observation
  in a sequence depends only on a previous unobserved, or hidden, state.~\cite{beal_infinite_2002}
  \begin{itemize}
    \item In the context of our simulations, the observations correspond to 
    the center of mass coordinates of the solutes versus time, and the states
    correspond to the dynamical behavior which give rise to those types
    of observations.
    \item Unfortunately, standard HMMs require the number of hidden states to be known
    a priori.
    \item One can partially overcome this by testing a range of numbers of 
    hidden states and determining which is the best representation of their
    data.
  \end{itemize}

  The infinite-state HMM overcomes this drawback by placing a hierarchical
  Dirichlet process (HDP) prior on the transition probabilities.
  \begin{itemize}
    \item Using some base probability distribution, H, a Dirichlet process 
    (DP) generates distributions over a countably infinite number of 
    probability measures:
    \begin{equation}
      G_0 = \sum_{k=1}^{\infty} \beta_k \delta_{\theta_k} ~~ \theta_k \sim H, \beta \sim GEM(\gamma)
    \end{equation}
    where the $\theta_k$ are values drawn from the base distribution and the
    weights $\beta_k$ come from a stick-breaking process parameterized by the concentration 
    parameter $\gamma$ (equivalently referred to as GEM($\gamma$)). 
    \item The concentration parameter expresses one's confidence in H relative to the posterior 
    and is closely related to the number of data observations.
    \item Each row, $G_j$, of the transition matrix is produced by drawing from a DP specified 
    using the $\beta$ vector as a discrete base distribution and a separate concentration
    parameter, $\alpha$.
    \begin{equation}
      G_j = \sum_{k=1}^{\infty} \pi_{jk} \delta_{\theta_k} ~~ \pi_j \sim DP(\alpha, \beta)
    \end{equation}
    \item This hierarchical specification ensures that the transition probabilities in 
    each row share the same support points \{$\theta_1$, ..., $\theta_k$\}.
    \item Once the model has converged only a finite number of states will have significant
    sampling.
  \end{itemize}

  \noindent We describe the dynamics of each state using a vector autoregressive (VAR) model. 
  \begin{itemize}
  	\item A VAR process is characterized by a vector of observations in a time series 
  	that are dependent on $r$ previous values of the time series vector, weighted by a
  	coefficient matrix $A_i$ in addition to a white noise term $\mathbf{e}_t$:
  	\begin{equation}
  	\mathbf{y}_t = \sum_{i=1}^r A_i\mathbf{y}_{t-i} + \mathbf{e}_t~~~~\mathbf{e}_t \sim N(0, \Sigma)
  	\end{equation}
  	\item We assumed multivariate Gaussian noise and limited our analysis to an 
  	autoregressive order of $r=1$.
  	\item We used a conjugate matrix-normal inverse-Wishart prior on parameters
  	$A$ and $\Sigma$ in order to analytically draw from the posterior.
  \end{itemize}

  Based on the VAR parameters and matrix of transition probabilities, we calculated
  the most likely sequence of hidden states.
  \begin{itemize}
    \item We repeated this process iteratively until we reached convergence % BJC: need to define convergence criteria
    \item Our python implementation of this process is heavily adapted from the MATLAB code
    of Fox et al.~\cite{fox_sticky_2007} 
    \item We refer the interested reader to much more extensive descriptions of 
    this process and its implementation. 
    ~\cite{beal_infinite_2002,teh_hierarchical_2006,van_gael_beam_2008,fox_nonparametric_2009,fox_bayesian_2010}
  \end{itemize}
  \break
  \section{Results and Discussion}
